---
title: "Summarize and Visualize Two Categorical Variables"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Step 1: Load the data set and define the colors.

```{r, echo = FALSE}
#eCornell Hex Codes: 
crimson = '#b31b1b' #Crimson
lightGray = '#cecece' #lightGray
darkGray = '#606366'
skyBlue = '#92b2c4' #skyblue
gold = '#fbb040' #gold
ecBlack = '#393f47' #ecBlack

titanic <- read.table("titanic.txt", header = TRUE)
SurvivedFactor <- factor(titanic$Survived, 
                         levels = c("0", "1"),
                         labels = c("No", "Yes"))
titanic$Survived <- SurvivedFactor
```

Q: Was the survival rate among male passengers higher than the survival rate among female passengers? 

We need to know four numbers: 
    1. Number of male passengers who survived 
    2. Number of female passengers who survived 
    3. Number of male passengers who did not survive 
    4. Number of female passengers who did not survive

## Step 2: Create a contingency table and calculate proportions of survivors grouped by sex.

Create a contingency table with the table() command that shows the number of individuals in each of the four groups of interest. 

Use the table() command on the whole data set.

```{r}
table(titanic$Sex, titanic$Survived)
```

Use the prop.table() command to calculate proportions of survivors.

```{r}
prop.table(table(titanic$Sex, titanic$Survived), 1)
# Setting second argument to 1 will give us proportions 
# within each row (% survivors among female, % survivors among male)

# Change this to 2 to calculate proportions
# within each column (e.g., % males among survivors)
```

## Step 3: Create a stacked barplot to show number of survivors grouped by sex.

Next, use the transpose command, t(), along with the barplot() command to create a stacked barplot that displays Sex on the x-axis. 

```{r}
tbl <- table(titanic$Sex,titanic$Survived)
tbl # Sex in rows, survival in columns
# The barplot function creates one bar for each column, so you'll need to transpose tbl before making a barplot
t(tbl) # Switch rows/columns to get one bar per gender
par(mar=2+c(5.1,4.1,4.1,2.1),cex.axis=1, cex.main=1.5, cex.lab=1.5)
barplot(t(tbl), 
        main="Survived?", cex.names = 1.5, 
        col = c(crimson, darkGray), # Bars will be crimson and darkGray
        legend.text=c("No","Yes"), # Create a legend
        args.legend=list(x="topleft",cex=1)) # Legend will be at the top left of the plot
```

## Step 4: Create a stacked barplot to show proportion of survivors grouped by sex.

Use the prop.table() command and  transpose the data to make a stacked barplot that shows proportions instead of numbers. 

```{r,echo=TRUE}
pt=prop.table(table(titanic$Sex,titanic$Survived),1)
t(pt) # Switch rows to columns
par(mar=2+c(5.1,4.1,4.1,2.1),cex.axis=1, cex.main=1.5, cex.lab=1.5)
barplot(t(pt), 
        main="Survived?", cex.names = 1.5,     
        col = c(crimson, darkGray),
        legend.text=c("No","Yes"),
        args.legend=list(x="topleft",cex=1))
```



