---
title: "Summarize and Visualize One Categorical Variable"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Step 1: Load the data set and define colors.

```{r, echo = FALSE}
#eCornell Hex Codes: 
crimson = '#b31b1b' #Crimson
lightGray = '#cecece' #lightGray
darkGray = '#606366'
skyBlue = '#92b2c4' #skyblue
gold = '#fbb040' #gold
ecBlack = '#393f47' #ecBlack

titanic <- read.table("titanic.txt", header = TRUE)
```

## Step 2: Change Survived to a factor.

Use the factor() function to change `titanic$Survived` into a vector called SurvivedFactor, then replace the original `titanic$Survived` variable with the new vector SurvivedFactor. 

```{r}
str(titanic)
SurvivedFactor <- factor(titanic$Survived, 
                         levels = c("0", "1"), 
                         labels = c("No", "Yes"))
titanic$Survived <- SurvivedFactor # Note that here we're replacing the Survived column with the SurvivedFactor column, instead of adding it to the data set as a new column

head(titanic, n = 10)
str(titanic)
```

Look at the first $11$ passengers in the titanic data: 

```{r, echo=FALSE}
head(titanic, n=11)
```

## Step 3: Construct a frequency table.

Use the table() command in R to construct a *frequency table*. Note that here we're only working with the first 11 passengers in the data set. 

```{r}
table(head(titanic$Survived, n=11)) # Make a table of the number of passengers that didn't survive and the number that did, from only the first 11 passengers. 
```

## Step 4: Make a relative frequency table.

Count *relative frequency* (proportion) of survivors and non-survivors in the data.

```{r}
# Manually check the proportion of survivors and non-survivors
6/(6+5) # Proportion that didn't survive
5/(6+5) # Proportion that survived
prop.table(table(head(titanic$Survived, n=11))) # Use the prop.table() function to do this automatically
```

## Step 5: Make frequency and relative frequency tables of all passenger survival.

Next, calculate the frequency and relative frequency for the entire titanic data set.

```{r}
tbl.titanic <- table(titanic$Survived)
tbl.titanic
prop.table(tbl.titanic)
```

## Step 6: Make a barplot of the survival of Titanic passengers.

Use the table you create of passenger survival to create a barplot with the command barplot(). 

```{r}
counts = table(titanic$Survived) # counts is a frequency table of the titanic$Survived column

par(mar=2+c(5.1,4.1,4.1,2.1)) # Set the margin around the plot
barplot(counts, # counts tells R the number of bars to make and the height of those bars
        main="Survival of Titanic Passengers", 
        col = crimson,
        ylab="Number of passengers", xlab="Survived?", 
        cex.axis=1.5, cex.main=1.5, 
        cex.names=1.5, cex.lab=1.5)
```

## Step 7: Make a barplot of the survival proportions of Titanic passengers.

Plot the proportion of survivors by dividing counts by the total number of passengers, then making a new barplot. 

```{r}
totnum = sum(counts)
par(mar=2+c(5.1,4.1,4.1,2.1))
barplot(counts / totnum, 
        main="Survival of Titanic Passengers", col = crimson,
        ylab="Proportion of passengers", xlab="Survived?",
        ylim=c(0,1),cex.axis=1.5, cex.main=1.5, 
        cex.names=1.5, cex.lab=1.5)
```

