---
title: "Measure the Error of a Fitted Line"
author: "eCornell"
date: "7/26/2021"
output:
  pdf_document: default
  html_document: default
---
Use this R Markdown file to understand how to draw a fit line through data and use error calculations to measure how well the line fits your data. 

## Step 1: Load the data and define colors.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(carData) # The Prestige data set is available in the carData library

data(Prestige) # Load the Prestige data set

# Exclude any observations that do not have an entry in the type column
Prestige = Prestige[!is.na(Prestige$type),] 

#eCornell Hex Codes: 
crimson = '#b31b1b' #Crimson
lightGray = '#cecece' #lightGray
darkGray = '#606366' #darkGray
skyBlue = '#92b2c4' #skyblue
gold = '#fbb040' #gold
ecBlack = '#393f47' #ecBlack

```

## Step 2: Plot the data, regression line, and errors.
Create a plot of prestige vs. education with the plot() function. Plot a line through it with the abline() function. Finally, use a for loop to plot the error lines between each point and its predicted value to help you visualize the difference between predicted and observed values. 

```{r}
# Create the plot: 
plot(prestige ~ education, data = Prestige, pch = 19, col = ecBlack,
     main = 'prestige = -10 + 5*education', ylim = c(15, 100))
# Plot the fit line. the argument a = -10 indicates the intercept is -10, b = 5 indicates the slope is 5:
abline(a = -10, b = 5, col = ecBlack, lwd = 2)
# for loop to draw vertical error lines:
for (i in 1:98){  
  lines(rep(Prestige$education[i], 2), 
        c(Prestige$prestige[i], -10+5*Prestige$education[i]),
        type = 'l', col = crimson, lwd = 1)
}
```

## Step 3: Assess how a line fits the data with MSE.
The following code indicates one way to calculate the Mean Square of Errors (MSE) in R. Use MSE to assess whether the line is a good fit to the data. MSE is a measurement that takes the average of the sum of the squared value of each error line.  

```{r}
# Add a column to the Prestige data set called fit1 that contains the predicted value of each point based on the prediction rule (education):
Prestige$fit1 = -10 + 5*Prestige$education
# Add a column to the Prestige data set called error1 that contains the difference between the observed and predicted values:
Prestige$error1 = Prestige$prestige - Prestige$fit1
# Calculate MSE by finding the average value of squared error: 
MSE1 = mean(Prestige$error1^2)
```

## Step 4: Visually compare line fits to data.
Compare the regression line prestige = -10 + 5*education with the fit of a different line: 
prestige = -10 + 7*education

```{r}
# Plot the line prestige = -10 + 7*education over the prestige vs. education scatterplot
plot(prestige ~ education, data = Prestige, pch = 19, col = ecBlack,
     main = 'prestige = -10 + 7*education', ylim = c(15, 100))
abline(a = -10, b = 7, col = ecBlack, lwd = 2)
# Add vertical error lines to the plot
for (i in 1:98){
  lines(rep(Prestige$education[i], 2), 
        c(Prestige$prestige[i], -10+7*Prestige$education[i]),
        type = 'l', col = crimson, lwd = 1)
}
```

## Step 5: Use MSE to compare line fits to data.
Use MSE to quantify how bad this fit is, relative to the other line. Higher MSE means higher error and, therefore, a worse fit. 

```{r}
# Find the MSE for this line:
Prestige$fit2 = -10 + 7*Prestige$education
Prestige$error2 = Prestige$prestige - Prestige$fit2
MSE2 = mean(Prestige$error2^2)
```