---
title: "Resampling With Bootstrap Methods"
author: "eCornell"
date: "7/21/2021"
output:
  pdf_document: default
  html_document: default
---
Scenario: Use this R Markdown file to help you create bootstrap samples of numerical data. This file outlines two examples: 
- one in which you want to compare the exam scores of two different course sections, and
- one in which you want to test a hypothesis about the median income in Canada based on the Prestige data set in R. You can start by creating bootstrap samples of the median income. 

# Step 1: Load the prestige data set and create colors.

```{r}
library(carData) # The Prestige data set is available in the carData library

data(Prestige)  # Load the Prestige data set

# Exclude any observations that do not have an entry in the type column:
Prestige = Prestige[!is.na(Prestige$type),] 

#eCornell Hex Codes: 
crimson = '#b31b1b' #Crimson
lightGray = '#cecece' #lightGray
darkGray = '#606366' #darkGray
skyBlue = '#92b2c4' #skyblue
gold = '#fbb040' #gold
ecBlack = '#393f47' #ecBlack
```

# Step 2: Bootstrap resampling on a small example.
Create the example data frame:
```{r}
dat = data.frame(Name = c("Alice", "Bob", "Catie", "Dave", "Eve"), Score = c(10, 20, 30, 40, 50))
dat # View the data frame
set.seed(1) #Set the seed for reproducibility
```

Then, to create bootstrap samples from this data set, you can use the `sample()` function with `replace = TRUE`.

```{r}
# This example data set will sample from the vector 1:5
1:5 
# Sample with replacement to create the boot.id vector
boot.id = sample(1:5, size = 5, replace = TRUE) 
boot.id
# Use the boot.id vector to create the bootstrap sample from your original data set
dat.boot = dat[boot.id,] 
# View the bootstrapped data set dat.boot
dat.boot 
```

# Step 3: Examine the prestige data set.

First, look at the profession types in the prestige data set.
```{r}
head(Prestige[Prestige$type == 'bc',])
head(Prestige[Prestige$type == 'wc',])
head(Prestige[Prestige$type == 'prof',])
```

Then, examine the histogram of incomes in this data set. Notice that the histogram is right skewed with some outliers, and has a median of 6035 CAD.

```{r}
# Create a histogram with a vertical line at the observed median income 
median(Prestige$income)
hist(Prestige$income, breaks = 20, freq = TRUE, col = ecBlack, border ='white',
     main = 'Histogram of Income', xlab = 'Income (CAD)')
# The abline() function creates a line on a plot. The argument v = x tells R to plot a vertical line at x.
abline(v=median(Prestige$income), col = crimson, lwd = 3)
# The legend command adds text to the plot.
legend('topright', legend = paste('median =', median(Prestige$income)), bty = 'n')
```

## Step 4: Use bootstrap resampling on the prestige data set.
Set the seed for reproducibility.
```{r}
set.seed(1)
```

Use `sample()` function to create a bootstrap data set. Run this code chunk several times to see the variation. 
```{r}
# Create bootstrap vector, boot.id
boot.id = sample(1:98, size = 98, replace = TRUE) 

# Create a new data frame, Prestige.boot, so that each row matches the rows sampled from the bootstrap vector.
Prestige.boot = Prestige[boot.id,] 

# Draw histogram of the bootstrapped data with the median plotted on it. 
hist(Prestige.boot$income, breaks = 20, freq = TRUE, col = ecBlack,
     border = 'white', main = 'Histogram of Income', xlab = 'Income (CAD)')
abline(v=median(Prestige$income), col = crimson, lwd = 3)
legend('topright', legend = paste('median =', median(Prestige.boot$income)), bty = 'n')

```

Notice that: 
- histograms of median income are changing across different bootstrap samples, and
- median incomes change from sample to sample, but all of the bootstrap data sets  maintain the overall shape of the histogram: right-skewed with some outliers.


