do ---
title: "Simulating_Under_Null_Hypothesis"
author: "eCornell"
date: "7/1/2021"
output: html_document
---

Scenario:
Use this R Markdown file to help you **simulate the null distribution of a sample statistic**. In this simulation, you will assume that the null hypothesis, which is that both new and old treatments have 50% chance of success (recovery), is true.

This example uses 'New' to represent a new drug, and 'Old' to represent an old drug in a clinical trial example. Both treatments have two possible outcomes — either the treatment works (is successful) or does not work (is not successful). 

## Step 1: Set up your code and create colors.

```{r}
# Set the seed so your results are reproducible. The first time you run code after using this command will be the same as any other time you run the same code.
set.seed(1)  

# Two possible outcomes — the individual recovered or did not recover.
outcome = c("Worked", "Did not Work") 

# eCornell Hex Codes: 
crimson = '#b31b1b'   # crimson
lightGray = '#cecece' # lightGray
darkGray = '#606366'  # darkGray
skyBlue = '#92b2c4'   # skyblue
gold = '#fbb040'      # gold
ecBlack = '#393f47'   # ecBlack
```

## Step 2: Simulate the outcomes of patients under the new treatment.
Create a vector of 52 values sampled from the outcome vector you created above. These values represent whether the treatment caused an individual to recover or not. 

```{r}
result_new = sample(outcome, # Vector from which to draw samples
    52, # Number of Samples
    replace = TRUE, # Sample with replacement
    prob = c(0.5, 0.5)) # The probabilities of success and failure for each patient
result_new # Display the result_new vector

p_new_sim = mean(result_new == "Worked") # Calculate the proportion of times the New treatment worked
p_new_sim # Display the proportion of times the new treatment worked
```

## Step 3: Simulate the outcomes of 52 patients under the old treatment. 
Use the same method you used in step 2, but select a sample of 51 individuals.
```{r}
result_old = sample(outcome, 51, replace = TRUE, prob = c(0.5, 0.5))
p_old_sim = mean(result_old == "Worked")
p_old_sim
```

## Step 4: Calculate the sample statistic.
The sample statistic is the difference in proportion of times that each treatment was successful.
```{r}
p_diff = p_new_sim - p_old_sim
p_diff
```

## Step 5: Run a simulation. 
Now, run your simulation by repeating steps 2-4 many, many times. 
```{r}
# The number of times you'll repeat the simulation
nsim = 100000 
# The vector in which to store the results of different simulations
store_p_diff = rep(0, nsim) 

# Create a for loop to tell R to run your simulation 100,000 times
for (i in 1:nsim){ 
  
   # Step 2:
    result_new = sample(outcome, 52, replace = TRUE, prob =     c(0.5, 0.5))
    p_new_sim = mean(result_new == "Worked")

   # Step 3:
    result_old = sample(outcome, 51, replace = TRUE, prob =     c(0.5, 0.5))
    p_old_sim = mean(result_old == "Worked")

   # Step 4:
    p_diff = p_new_sim - p_old_sim
    store_p_diff[i] = p_diff
 
} # Close the for loop
```

## Step 6: Visualize the distribution of sample statistics.
Visualize the distribution of sample statistics by creating a histogram.

```{r}
hist(store_p_diff, breaks = 40, freq = FALSE, col = ecBlack,
     main = 'Histogram of Sample Differences (New - Old)',
     xlab = 'Difference in Success Rates (New - Old)')
```

## Step 7: Interpret the simulation results.
Calculate the average sample statistic across all 100,000 simulation results. 

```{r}
mean(store_p_diff)
```
We see that this histogram of sample statistics is centered around 0, but we also see that the difference is as large as 10% or higher in many simulations.

Calculate the number of times the sample statistic was more than +/- 10%. This means the New and Old treatments had success rates that were at least 10% different.
```{r}
mean(store_p_diff > 0.1 | store_p_diff < -0.1)
```

In about 32% of simulations (roughly 1 in 3 simulations), new and old treatment and new treatment had at least a 10% different success rate. 

Calculate the standard deviation of the sample statistics.

```{r}
sd(store_p_diff)
```

We see that on average the difference between the success rates of new and old treatment varied about 9.8% around 0, even though in truth (population) their success rates were the same.
