---
title: "Convey Y About X Data Set"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Set the theme and load your data.

```{r, warning = FALSE, message = FALSE}
# 1) Set up the theme used in these videos:
library(ggplot2)

theme <-  theme(plot.margin = margin(5, 5, 5, 5, "pt"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          panel.border = element_rect(colour = "#393f47", fill = NA, size = 2),

          axis.text = element_text(size = 20),
          axis.title.x = element_text(size = 24),
          axis.title.y = element_text(size = 24),
          plot.title = element_text(face = "bold", size = 30))

ourTheme <- list(theme, scale_color_manual(values = c('#393f47',
       '#b31b1b', '#fbb040', '#92b2c4', '#cecece')))

# 2) Load data from the National Oceanic and Atmospheric Administration's 
# Atlantic hurricane database and convert variables to factors where necessary.

library(tidyverse)

# read in the storm data:
storms <- read.csv("storms.csv")

# set the storm category to be a factor:
storms$Category <- factor(storms$Category, levels = -1:5) 

# set the measurement date/time to be a factor:
storms$Date <- factor(storms$Date, levels = unique(storms$Date)) 

# View(storms) # Look at the storms data

# 3) Filter the data to only use observations for Hurricane Sandy

sandy <- storms %>% filter(Name == "Sandy")
```

## Step 2: Make a scatterplot of TS_diameter versus air pressure for Hurricane Sandy.
When you don't specify a color, all the points on the plot will be black: 

```{r basic scatterplot}
# basic scatterplot (all points black)
ggplot(data = sandy, aes(x = TS_diameter, y = Pressure)) + 
        geom_point(size = 4) + 
        ourTheme
```

## Step 3: Adjust point color to reflect storm category.
When you adjust points within the aes() function, the change will affect every point. Here, setting color = Category means point color depends on the storm category:

```{r changing color of points}
# color of points depends on storm category
ggplot(data = sandy, aes(x = TS_diameter, y = Pressure, color = Category)) + 
        geom_point(size = 4) + 
        ourTheme
```

## Step 4: Change the point shapes to reflect category.
When you adjust points within the aes() function, the change will affect every point. Here, setting shape = Category means point shape depends on the storm category:

```{r changing shape of points}
# shape of points depends on storm category
ggplot(data = sandy, aes(x = TS_diameter, y = Pressure, shape = Category)) + 
        geom_point(size = 4) + 
        ourTheme
```

## Step 5: Change the point transparency to reflect wind speed.
When you adjust points within the aes() function, the change will affect every point. Here, setting alpha = Wind means point transparency depends on the wind speed:

```{r changing transparency of points}
# transparency of points depends on wind speed
ggplot(data = sandy, aes(x = TS_diameter, y = Pressure, alpha = Wind)) + 
        geom_point(size = 4) + 
        ourTheme
```

## Step 6: Change the point size to reflect wind speed.

```{r changing size of points}
# size of points depends on wind speed
ggplot(data = sandy, aes(x = TS_diameter, y = Pressure, size = Wind)) + 
        geom_point() + 
        ourTheme
```

## Step 7: Combine aesthetic adjustments.
You can adjust multiple aesthetics at once. Here, setting color = Category and size = Wind means point color depends on Category and point size depends on the wind speed:

```{r multiple aesthetic adjustments}
# color of points depends on storm category, size of points depends on wind speed
ggplot(data = sandy, aes(x = TS_diameter, y = Pressure, color = Category, size = Wind)) + 
        geom_point() + 
        ourTheme
```
