---
title: "Change Geometric Plot Features"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Set the theme and load your data.

```{r, warning = FALSE, message = FALSE}
# 1) Set up the theme used in these videos:
library(ggplot2)

theme <-  theme(plot.margin = margin(5, 5, 5, 5, "pt"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          panel.border = element_rect(colour = "#393f47", fill = NA, size = 2),
          axis.text = element_text(size = 20),
          axis.title.x = element_text(size = 24),
          axis.title.y = element_text(size = 24),
          plot.title = element_text(face = "bold", size = 30))

ourTheme <- list(theme,

scale_color_manual(values = c('#393f47', '#b31b1b', '#fbb040', '#92b2c4')),
        scale_fill_manual(values = c('#b31b1b', '#cecece', '#393f47', '#92b2c4', '#fbb040')))

# 2) Load data from the National Oceanic and Atmospheric Administration's 
# Atlantic hurricane database and convert variables to factors where necessary:

library(tidyverse)

# read in the storm data:
storms <- read.csv("storms.csv")

# set the storm category to be a factor:
storms$Category <- factor(storms$Category, levels = -1:5) 

# set the measurement date/time to be a factor:
storms$Date <- factor(storms$Date, levels = unique(storms$Date)) 

# look at the storm data
view(storms)

# 3) Filter the data to only use observations for Hurricane Sandy:

sandy <- storms %>% filter(Name == "Sandy")

# 4) Create a data set with observations only for Hurricanes Katrina, Sandy, and Wilma:

sampleStorms <- storms %>% filter(Name %in% c("Katrina", "Sandy", "Wilma"))

head(sampleStorms)
```

## Step 2: Make a barplot of the number of observations for each storm.
Using geom_bar() instead of geom_point() makes the geometry of this plot a bar. Since this is a barplot of number of observations, there is no need to specify a y variable in the aesthetic function because R does this calculation for you. The color of the bars is specified by the argument fill:

```{r barplot}
# use geom_bar() for barplots
ggplot(sampleStorms, aes(x = Name)) + 
        geom_bar(fill = '#b31b1b') + 
        ourTheme
```

## Step 3: Create a stacked barplot. 
To create a stacked barplot, you can adjust the fill argument within the aes() function. The groups within the barplot are created by setting fill = Category:

```{r stacked barplot}
ggplot(sampleStorms, aes(x = Name, fill = Category)) +
        geom_bar() + 
        ourTheme
```

## Step 4: Create a side-by-side barplot.
Adding the argument position = "dodge" within the geom_bar() function makes the barplot a side-by-side barplot, instead of a stacked barplot:

```{r side-by-side barplot}
ggplot(sampleStorms, aes(x = Name, fill = Category)) + 
        geom_bar(position = "dodge") + 
        ourTheme
```

## Step 5: Make a line plot of wind speed over time for Hurricane Sandy.
Use the geom_line() function to make a line plot. Make sure to set group = 1 within the aes() function since the x-axis variable is a factor:

```{r line plot}
ggplot(sandy, aes(x = Date, y = Wind, group = 1)) + 
        geom_line(size = 3) + 
        ourTheme
```
