---
title: "Combine_Transformations_Efficiently.Rmd"
author: "eCornell"
date: "3/15/2021"
output:
  pdf_document: default
  html_document: default
---
Use these libraries and this data set for all questions:

```{r setup, include=FALSE}
library(tidyverse)
library(ellipsis)
Credit <- read.csv("credit_cards.csv")
```

*CQ1:* 
Use the pipe operator, %>%, to combine transformations to create a data frame called CQ1 that satisfies the following criteria:
(a) Includes only those individuals with two or more credit cards 
(b) Only displays the variables Income and Cards
```{r, echo=FALSE}
Credit <- read.csv("credit_cards.csv")
CQ1 = Credit %>% filter(Cards >= 2) %>% select(Income,Cards)
```

*CQ2:*
Use the pipe operator, %>%, to combine transformations to create a data frame called CQ2 that satisfies the following criteria:
(a) Does not include the ID variable 
(b) Contains a variable named HigherEd that is TRUE if the person has more than 13 years of education and FALSE otherwise 
(c) Contains a variable named Unused that gives the unspent portion of the credit limit (i.e., the limit minus the balance)
```{r, echo=FALSE}
Credit <- read.csv("credit_cards.csv")
CQ2 <- Credit %>% 
  select(-c(ID)) %>% 
  mutate(HigherEd = Education>13) %>% 
  mutate(Unused = Limit-Balance)
```

*CQ3*
Use the pipe operator, %>%, to combine transformations to create a data frame called CQ3 that satisfies the following criteria:
(a) Contains a variable named Unused that gives the unspent portion of the credit limit (i.e., the limit less the balance) 
(b) Contains only those individuals for whom the unspent portion of their credit limit is at least $1,000
```{r, echo=FALSE}
Credit <- read.csv("credit_cards.csv")
CQ3 <- Credit %>% mutate(Unused=Limit-Balance) %>% filter(Unused >=1000)
```