---
title: "Create and Alter Date-Time Objects From Strings"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Step 1: Load your data.

```{r load packages and data set, warning=FALSE, message=FALSE}
library(tidyverse)
calls <- read.csv("calls.csv", stringsAsFactors = FALSE)

# Prepare data:

# Initialize vectors to store each of the new variables
address <- c()
town <- c()
dt <- c()

for(i in 1:nrow(calls)) { # loop over emergency calls
  
  # get the description of the i^th call
  callI <- calls[i, "desc"] 
  
  # split the description text based on ";" --> gives a matrix of substrings
  splitCallDesc <- str_split(callI, ";", simplify = TRUE) 
  
  # store the street address, town, and date/time
  address[i] <- splitCallDesc[1]
  town[i] <- splitCallDesc[2]
  dt[i] <- splitCallDesc[3]
  
}

# add the new variables to the data frame
calls$address <- address
calls$towns <- town
calls$dt <- dt

# For plot:
theme <-  theme(plot.margin = margin(5, 5, 5, 5, "pt"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        panel.border = element_rect(colour = "#393f47", fill = NA, size = 2),
        axis.text = element_text(size = 30),
        axis.title.x = element_text(size = 30),
        axis.title.y = element_text(size = 30),
        plot.title = element_text(face = "bold", size = 30))
ourTheme <- list(theme, scale_size_manual(values = c(1.5)))
```

## Step 2: Examine your data.
Look at the first few rows of the data. Notice that the dt variable contains date-times.

```{r looking at date-time variable}
head(calls$dt)
```

## Step 3: Load the lubridate package.
This package is useful for working with date-times.  

```{r loading lubridate, warning = FALSE, message = FALSE}
library(lubridate)
```

## Step 4: Convert date-times from strings to POSIXct objects.

```{r creating a POSIXct object}
# convert date-times from strings to POSIXct objects
# need to specify the order in which the components appear -- in this case, year-month-day-hour-minute-second
calls$dtBetter <- ymd_hms(calls$dt)
head(calls$dtBetter)
```

## Step 5: Set the time zone to follow New York.

```{r creating a POSIXct object with time zone specified}
# convert date-times from strings to POSIX objects, but with the time zone specified
temp <- ymd_hms(calls$dt, tz = "America/New_York")
head(temp)
```

## Step 6: Extract components from the date-time of the first call.

```{r extracting components of date-time}
# Date and time of first call
firstCall <- calls$dtBetter[1]
firstCall

# Pull out components
year(firstCall)
month(firstCall)
day(firstCall)
hour(firstCall)
minute(firstCall)
second(firstCall)
```

## Step 7: Make a new variable giving the hour of each 911 call.

```{r make hour variable}
# create a new variable called "hour" that gives the hour (0-23) of the emergency call
calls$hour <- hour(calls$dtBetter)
head(calls)
```

## Step 8: Create a barplot that illustrates how call volume changes over the course of the day.

```{r barplot of number of calls}
# make a barplot of the number of calls in each hour
ggplot(calls, aes(x = hour)) + geom_bar(fill = "#b31b1b") + ourTheme
```
