---
title: "Extract_Information_Quiz.Rmd"
author: "eCornell"
date: "3/16/2021"
output:
  pdf_document: default
  html_document: default
---
Use these libraries and data set in the following code chunk as the starting point to answer all of the questions below.

```{r, include=FALSE}
library(tidyverse)
library(lubridate)
library(utf8)
calls <- read.csv("calls_CQ1_Q2.csv", stringsAsFactors = FALSE)
calls$address<- str_replace(calls$address, " & ", " @ ")

```

*CQ1:*
Write code to extract the specific reasons for all cardiac-related calls (i.e., extract any values of calls$reason that have the word "CARDIAC" in them), and save this subset as a vector called reasons.
```{r, echo=FALSE}
indices <- str_which(calls$reason, "CARDIAC")
reasons = calls[indices,]$reason
```

*CQ2:*
Earlier we saw that some of the addresses specify an intersection (e.g., MAIN ST @ BUTTONWOOD ST), but others only give one street (e.g., W MAIN ST). We might want to count how many addresses only contain one street since these are clearly less detailed. Edit the str_which() function below to do so.
```{r, echo=FALSE}
temp <- str_which(calls$address, " @ ") # how many locations have an intersection
                    #(remember that "@" is used to denote an intersection)
length(temp) # how many calls have an intersection
```
