---
title: "Extract Particular Information From Strings Where It Exists"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Load and process your data.
Load the calls data set and process it to split the address, town, and date-time variables: 

```{r load packages and data set, warning = FALSE, message = FALSE}
library(tidyverse)
calls <- read.csv("calls.csv", stringsAsFactors = FALSE)

# Initialize vectors to store each of the new variables
address <- c()
town <- c()
dt <- c()

for(i in 1:nrow(calls)) { # loop over emergency calls
  
  # get the description of the i^th call
  callI <- calls[i, "desc"] 
  
  # split the description text based on ";" --> gives a matrix of substrings
  splitCallDesc <- str_split(callI, ";", simplify = TRUE) 
  
  # store the street address, town, and date/time
  address[i] <- splitCallDesc[1]
  town[i] <- splitCallDesc[2]
  dt[i] <- splitCallDesc[3]
}

# add the new variables to the data frame
calls$address <- address
calls$towns <- town
calls$dt <- dt
```

## Step 2: Find rows that contain the string "VEHICLE".
Use str_which() to find the indices of the vehicle-related emergency calls, then count how many of them there were with the length() function.

```{r finding whether match exists with str_which()}
# find which strings in the vector calls$title contain the pattern "VEHICLE"
indices <- str_which(calls$title, "VEHICLE")
head(indices)
```

In this case, we wanted to find a word, but this function can find any string that exactly matches the second argument. Then, we can find those rows in our data set:

```{r finding number of strings for which match exists}
calls[indices,]
```

## Step 3: Find strings that contain the string "VEHICLE".
To find all the strings in which a word or pattern appears, use the str_subset() function: 
```{r extracting the match with str_subset()}
temp <- str_subset(calls$title, "VEHICLE")
head(temp)
```

