---
title: "Filter_Select_Quiz.Rmd"
author: "eCornell"
date: "8/7/2024"
output:
  pdf_document: default
  html_document: default
---
Use these libraries and this data set for all questions:

```{r setup}
library(tidyverse)
Credit <- read.csv("credit_cards.csv")
```

*CQ1:* 
Use the filter() function to keep only those individuals who are under 50 years old in the Credit data set. Assign the new data set to an object called CQ1.
```{r}
Credit <- read.csv("credit_cards.csv")
CQ1 <- filter(Credit, Age < 50)
head(CQ1)
```

*CQ2:* 
Use the filter() function to keep only those individuals who have more than one credit card and are not a student. Assign the new data set to an object called CQ2.
```{r}
Credit <- read.csv("credit_cards.csv")
CQ2 <- filter(Credit, Student == 'No' & Cards >= 2)
head(CQ2)
```

*CQ3:* 
Use the filter() function to keep only those individuals who have a credit rating above 350 or an income greater than $45,000 (recall that income is measured in units of $1,000). Assign the new data set to an object called CQ3.
```{r}
Credit <- read.csv("credit_cards.csv")
CQ3 <- filter(Credit, Rating > 350 | Income > 45)
head(CQ3)
```

*CQ4:*
Use the select() function to remove the Married variable from the Credit data set. Assign the new data set to an object called CQ4.
```{r}
Credit <- read.csv("credit_cards.csv")
CQ4 <- select(Credit, -c(Married))
head(CQ4)
```

*CQ5:*
Use the select() function to keep only the Limit, Cards, and Balance variables in the Credit data set. Assign the new data set to an object called CQ5.
```{r}
Credit <- read.csv("credit_cards.csv")
CQ5 <- select(Credit, c(Limit,Cards,Balance))
head(CQ5)
```