---
title: "Filtering Data"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Load your data.

```{r loading data and library,warning=FALSE,message=FALSE}
library(tidyverse)
Credit <- read.csv("credit_cards.csv")
```

## Step 2: Filter data with the filter() function and the logical operator "==".
Recall that the filter() function takes two arguments: the argument to specify the data set and the argument for how you'll filter the observations in your data.

Here, you'll use the filter() function to select only students by using the logical operator for "is equal to," which is "==": 

```{r}
filter(Credit, Student == "Yes")
```

## Step 3: Filter data with the logical operator ">".
Next, filter your data to select only those credit card holders that have a balance greater than zero:

```{r}
filter(Credit, Balance > 0)
```

## Step 4: Filter data with multiple logical operators.
Finally, filter your data by multiple attributes: 

```{r}
filter(Credit, Student == "Yes" & Balance > 0)

filter(Credit, Student == "Yes" | Education > 13)
```
 
## Step 5: Deselect unwanted variables with the select() function.
The select() function also takes two arguments: the data set and the variables you would like to select or deselect. Here, you'll deselect ID from the Credit data set:

```{r}
select(Credit, -ID)
```
 
## Step 6: Select or deselect a set of variables.
To deselect a set of variables, use the c() function to combine the variables, then add a minus sign in front of the set. To select a set, don't add a minus sign in front of the set:

```{r}
select(Credit, -c(ID, Income))

select(Credit, c(Income, Limit, Rating))
```
 