---
title: "Modify_Strings_Quiz.Rmd"
author: "eCornell"
date: "3/16/2021"
output:
  pdf_document: default
  html_document: default
---
Use these libraries and this data set as the starting point to answer all of the questions below.

```{r, include=FALSE}
library(tidyverse)
library(lubridate)
library(utf8)

```

*CQ1:*
Use the code in the code chunk below as a starting point, then write code so that intersections in the Address variable, currently denoted with a `&`, are expressed with the `@` symbol instead; e.g., MAIN ST @ BUTTONWOOD ST. Be sure to keep track of any white space that is needed. 
```{r, echo=FALSE}
calls<-read.csv("calls_CQ1_Q2.csv", stringsAsFactors = F)
calls$address <- str_replace(calls$address, " & ", " @ ")
```

*CQ2:*
Use the code in the code chunk below as a starting point, then write code to remove excess white space around the values of the reason, towns, time, and date variables.
```{r, echo=FALSE}
calls<-read.csv("calls_CQ1_Q2.csv", stringsAsFactors = F)

calls$reason <- str_trim(calls$reason)
calls$towns <- str_trim(calls$towns)
calls$time <- str_trim(calls$time)
calls$date <- str_trim(calls$date)

```
*CQ3:* 
Use the code in the code chunk below as a starting point, then write code to convert all of the values of reason to lowercase.
```{r, echo=FALSE}
calls<-read.csv("calls_CQ1_Q2.csv", stringsAsFactors = F)
calls$reason <- tolower(calls$reason)
```