---
title: "Use pivot_wider and pivot_longer"
author: "eCornell"
date: "10/8/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Load your data.

```{r}
library(tidyverse)
langDivLengthen <- read.csv("language_diversity_pl.csv") 
head(langDivLengthen)
```
```{r}
langDivWiden <- read.csv("language_diversity_pw.csv") 
head(langDivWiden)
```



## Step 2: Lengthen a data set.
Use the pivot_longer() function to lengthen a data set. pivot_longer() takes the arguments data for the data set, cols for the columns whose names are values of a variable, names_to for what to call the new column, and values_to for what to call the column into which the cells of the data frame will be moved.

Note that even though this code lengthens your data set, it still needs additional manipulation before it is tidy. 
```{r}
langDiv1 <- pivot_longer(data = langDivLengthen, 
                cols = -Measurement, 
                names_to = "Country",
                values_to = "Value")
head(langDiv1)
```

## Step 3: Widen a data set.
Use the pivot_wider() function to widen a data set. pivot_wider() takes the arguments data for the data set, names_from for the column of data that contains the names of the new columns, and values_from for the column of data that contains the values for the new columns.

```{r}
langDiv2 <- pivot_wider(data = langDivWiden,
            names_from = Measurement,
            values_from = Value)

head(langDiv2)
```