---
title: "Splitting Strings Into Useable Pieces"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Load and examine your data.
```{r load packages and data set, warning = FALSE, message = FALSE}
# Load the calls data set: 
library(tidyverse)
calls <- read.csv("calls.csv", stringsAsFactors = FALSE)

# Then, examine it to see how it is organized: 
head(calls)
```

## Step 2: Split the description of the first call in the data set.
Extract the description of the first call and assign it to the variable temp:

```{r}
temp <- calls$desc[1]
temp
```
Since the description uses semicolons to split each fragment, that is what you'll use to separate your data when you use the str_split() function:
```{r}
head(calls)
```

Split the first call description into components.

```{r split a single string with str_split()}
str_split(temp, ";", simplify = TRUE)
```

## Step 3: Split all call descriptions with a for loop.
To separate the descriptions of each call in the whole data set quickly, you'll want to write a for loop. The comments within the following for loop describe what each step of the loop does.   

```{r loop to split multiple strings with str_split()}
# Initialize vectors to store each of the new variables:

address <- c()
town <- c()
dt <- c()

for(i in 1:nrow(calls)) { # loop over emergency calls
  
  # get the description of the i^th call
  callI <- calls[i, "desc"] 
  
  # split the description text based on ";" --> gives a matrix of substrings
  splitCallDesc <- str_split(callI, ";", simplify = TRUE) 
  
  # store the street address, town, and date/time
  address[i] <- splitCallDesc[1]
  town[i] <- splitCallDesc[2]
  dt[i] <- splitCallDesc[3]
  
}
```

## Step 4: Store the new vectors in the calls data frame.
Now that you've created the address, town, and date-time (dt) vectors, you can add them to the data frame as new variables:

```{r add new variables to the data frame}
calls$address <- address
calls$towns <- town
calls$dt <- dt
```

## Step 5: Examine the new data set.
```{r}
head(calls)
```
