---
title: "Summarize_Group_Quiz.Rmd"
author: "eCornell"
date: "3/15/2021"
output:
  pdf_document: default
  html_document: default
---
Use these libraries and this data set for all questions:

```{r setup, include=FALSE}
library(tidyverse)
library(ellipsis)
Credit <- read.csv("credit_cards.csv")
```

*CQ1:* 
Complete the rest of the starter code in the code chunk below to determine which group of individuals — those who are 50 or under versus those who are over 50 — have a higher median credit limit. Call your new summary statistic MedianLimit and assign your new data frame to the variable CQ1.
```{r, echo=FALSE}
Credit <- read.csv("credit_cards.csv")

#Starter code: 
CQ1<- Credit %>%
         mutate(Over50 = Age > 50) %>% 
         group_by(Over50) %>%
         summarise(MedianLimit=median(Limit))
```

*CQ2:* 
Use the count function, n(), to find the number of individuals who are: 
(a) Not a student and not married
(b) Not a student and married 
(c) A student and not married 
(d) A student and married
Call your new summary statistic NumInGroup and assign your new data frame to the variable CQ2.

```{r, echo=FALSE}
Credit <- read.csv("credit_cards.csv")
CQ2 <- Credit %>% 
  group_by(Student,Married) %>%
  summarize(NumInGroup = n())
```